lifespan_eda

### 2025-01-27 eda/lifespan_eda.Rmd
cleaning up data filtering and curation

Plot full samples for supplemental figures looking at imbalanced demographics across sites

@@

```{r}
#drop repeat scans
cn.df.unique <- cn.df %>%
  drop_na(c(age_days, participant, study)) %>%
  arrange(age_days) %>%
  group_by(participant) %>%
  slice(1) %>%
  ungroup() %>%
  mutate(sexMale = as.numeric(case_when(sex == "Male" ~ 1,
                                        sex == "Female" ~ 0,
                                        TRUE ~ NA)),
         sex.age = sexMale*age_days,
         log_age = log(age_days, base=10))

nrow(cn.df.unique) == n_unique(cn.df.unique$participant)
nrow(cn.df) - nrow(cn.df.unique)
```

@@

## Visualize Features
plot global brain features for basic qc - check again post-combat

@@

```{r}
cn.df.noinf <- cn.df.unique %>%
  dplyr::filter(age_days >= (365.25*3+280)) %>%
  mutate(study = droplevels(study),
         site = droplevels(site))
nrow(cn.df.unique) - nrow(cn.df.noinf)
n_unique(cn.df.unique$study) - n_unique(cn.df.noinf$study)
min(cn.df.noinf$age_days)
```

Visualizing euler distributions
```{r}
cn.df.noinf %>%
  ggplot() +
  geom_boxplot(aes(x=study, y=surfholes))
```

@@

```{r}
cn.df.eulerfilt <- cn.df.noinf %>%
  dplyr::filter(!is.na(surfholes)) #drops 2 studies w missing euler data

@@

Removed `r nrow(cn.df.noinf) - nrow(cn.df.eulerfilt2)` subjects, `r nrow(cn.df.eulerfilt2)` remain

@@

```{r}
#sample size, sex
length(unique(cn.df_mad_filt$participant))
nrow(cn.df_mad_filt) == length(unique(cn.df_mad_filt$participant))

@@

###Plot

@@

```{r}
fig5_AB <- ggarrange(fig5A, full_rain,
          nrow=1, ncol=2, 
          #legend = "none",
          widths=c(1.8, 1),
          labels = "AUTO")
fig5_AB
saveRDS(fig5_AB, file="../figs/figure_5_AB.rds")
```

>>>>

#Visualize
Plot full samples for supplemental figures looking at imbalanced demographics across sites

@@

```{r}
cn.df.noinf <- cn.df %>%
  dplyr::filter(age_days >= (365.25*3+280)) %>%
  mutate(study = droplevels(study),
         site = droplevels(site))
nrow(cn.df) - nrow(cn.df.noinf)
n_unique(cn.df$study) - n_unique(cn.df.noinf$study)
min(cn.df.noinf$age_days)
```

```{r}
set.seed(1212)
#drop repeat scans
cn.df.unique <- cn.df.noinf %>%
  drop_na(c(age_days, INDEX.ID, study)) %>%
  arrange(age_days) %>%
  group_by(INDEX.ID) %>%
  slice_sample(n=1) %>%
  ungroup() %>%
  mutate(sexMale = as.numeric(case_when(sex == "Male" ~ 1,
                                        sex == "Female" ~ 0,
                                        TRUE ~ NA)),
         sex.age = sexMale*age_days,
         log_age = log(age_days, base=10))

nrow(cn.df.unique) == n_unique(cn.df.unique$INDEX.ID)
nrow(cn.df.noinf) - nrow(cn.df.unique) #n scans
n_unique(cn.df.noinf$INDEX.ID[duplicated(cn.df.noinf$INDEX.ID)]) #n subjects
```

@@

plot global brain features for basic qc - check again post-combat

@@

@@

```{r}
cn.df.eulerfilt <- cn.df.unique %>%
  dplyr::filter(!is.na(surfholes)) #drops 2 studies w missing euler data

@@

Removed `r nrow(cn.df.unique) - nrow(cn.df.eulerfilt2)` subjects, `r nrow(cn.df.eulerfilt2)` remain

@@

```{r}
#sample size, sex
length(unique(cn.df_mad_filt$INDEX.ID))
stopifnot(nrow(cn.df_mad_filt) == length(unique(cn.df_mad_filt$INDEX.ID)))

@@

##Plot

@@

```{r}
fig5_AB <- ggarrange(fig5A, full_rain,
          nrow=1, ncol=2, 
          #legend = "none",
          widths=c(1.8, 1),
          labels = "AUTO")
fig5_AB
#saveRDS(fig5_AB, file="../figs/figure_5_AB.rds")
ggsave("../figs/fig5_ab.jpeg", fig5_AB, width=18, height=10, units="in")
```
